Global effects in locally optimal DPG test functions for convection-diffusion problems

### 2012-12-04 docs/Jesse/GlobalLocalTesting/paper.tex

\newcommand{\mb}[1]{\mathbf{#1}}
\newcommand{\mbb}[1]{\mathbb{#1}}
\newcommand{\mc}[1]{\mathcal{#1}}
\newcommand{\nor}[1]{\left\| #1 \right\|}
\newcommand{\snor}[1]{\left| #1 \right|}
\newcommand{\LRp}[1]{\left( #1 \right)}
\newcommand{\LRs}[1]{\left[ #1 \right]}
\newcommand{\LRa}[1]{\left\langle #1 \right\rangle}
\newcommand{\LRc}[1]{\left\{ #1 \right\}}
\newcommand{\tanbui}[2]{\textcolor{blue}{\sout{#1}} \textcolor{red}{#2}}
\newcommand{\Grad} {\ensuremath{\nabla}}
\newcommand{\Div} {\ensuremath{\nabla\cdot}}
\newcommand{\Nel} {\ensuremath{{N^\text{el}}}}
\newcommand{\jump}[1] {\ensuremath{\LRs{\![#1]\!}}}
\newcommand{\uh}{\widehat{u}}
\newcommand{\fnh}{\widehat{f}_n}
\renewcommand{\L}{L^2\LRp{\Omega}}
\newcommand{\pO}{\partial\Omega}
\newcommand{\Gh}{\Gamma_h}
\newcommand{\Gm}{\Gamma_{-}}
\newcommand{\Gp}{\Gamma_{+}}
\newcommand{\Go}{\Gamma_0}
\newcommand{\Oh}{\Omega_h}

@@

\author{Jesse Chan\textsuperscript{a}, Norbert Heuer\textsuperscript{b}, Tan Bui-Thanh\textsuperscript{a}, and Leszek Demkowicz\textsuperscript{a}}

%\title{$\epsilon$-explicit analysis of DPG for convection-dominated diffusion}
\title{Globally and locally optimal test functions}
\date{}
\begin{document}

%\tableofcontents
\maketitle
\begin{center}
\textsuperscript{a} Institute for Computational Engineering and Sciences, \\University of Texas at Austin, \\Austin, TX 78712, USA\\
\end{center}

\begin{center}
\textsuperscript{b} Facultad de Matem\'aticas, \\Pontificia Universidad Cat\'olica de Chile,\\
Avenida Vicu\~na Mackenna 4860, Santiago, Chile
\end{center}

\section{The ultra-weak variational formulation}

@@

Assuming boundedness below of $A$, as the extra $L^2$ term disappears, we get $L^2$ optimality. The ultra-weak variational form gives
\begin{align*}
b(u_h,v) = \LRp{u_h,A^*v} + \LRa{\widehat{u},v} = \text{load and boundary conds} = b(u,A^*v)
\end{align*}
If we use for our test functions solutions of the adjoint equation $A^*v = u$ (with appropriate adjoint boundary conditions), $u_h$ is the best $L^2$ projection of the exact solution $u$ onto $U_h$.

\section{Test norms for the convection-diffusion problem}

Introduce the graph norm here.

\subsection{Local and global optimal test functions}

Note that globally optimal test functions only produce boundary layers at the global inflow boundary.

Show representation of global test functions in local test functions using Jay's proof using orthogonal subspaces and flux test functions.

>>>>

\newcommand{\mb}[1]{\mathbf{#1}}
\newcommand{\mbb}[1]{\mathbb{#1}}
\newcommand{\mc}[1]{\mathcal{#1}}
\newcommand{\nor}[1]{\left\| #1 \right\|}
\newcommand{\snor}[1]{\left| #1 \right|}
\newcommand{\LRp}[1]{\left( #1 \right)}
\newcommand{\LRs}[1]{\left[ #1 \right]}
\newcommand{\LRa}[1]{\left\langle #1 \right\rangle}
\newcommand{\LRc}[1]{\left\{ #1 \right\}}
\newcommand{\tanbui}[2]{\textcolor{blue}{\sout{#1}} \textcolor{red}{#2}}
\newcommand{\Grad} {\ensuremath{\nabla}}
\newcommand{\Div} {\ensuremath{\nabla\cdot}}
\newcommand{\Nel} {\ensuremath{{N^\text{el}}}}
\newcommand{\jump}[1] {\ensuremath{\LRs{\!\left[#1\right]\!}}}
\newcommand{\uh}{\widehat{u}}
\newcommand{\fnh}{\widehat{f}_n}
\renewcommand{\L}{L^2\LRp{\Omega}}
\newcommand{\pO}{\partial\Omega}
\newcommand{\Gh}{\Gamma_h}
\newcommand{\Gm}{\Gamma_{-}}
\newcommand{\Gp}{\Gamma_{+}}
\newcommand{\Go}{\Gamma_0}
\newcommand{\Oh}{\Omega_h}

@@

\newtheorem{remark}{Remark}

\author{Jesse Chan\textsuperscript{a}, Jay Gopalakrishnan\textsuperscript{b}, and Leszek Demkowicz\textsuperscript{a}}

%\title{$\epsilon$-explicit analysis of DPG for convection-dominated diffusion}
\title{Global effects in locally optimal DPG test functions for convection-diffusion problems}
\date{}
\begin{document}

\maketitle
\begin{center}
\textsuperscript{a} Institute for Computational Engineering and Sciences, \\University of Texas at Austin, \\Austin, TX 78712, USA\\
\end{center}

\begin{center}
\textsuperscript{b} Department of Mathematics\\Portland State University, \\
Portland, OR 97207, USA
%\textsuperscript{b} Facultad de Matem\'aticas, \\Pontificia Universidad Cat\'olica de Chile,\\
%Avenida Vicu\~na Mackenna 4860, Santiago, Chile
\end{center}

\tableofcontents

\section{Introduction to DPG}

The Discontinuous Petrov-Galerkin method was originally introduced by Demkowicz and Gopalakrishnan in 2008 method as a way to attain automatic inf-sup stability for any well-posed variational problem through the construction optimal test functions \cite{DPG2}.  The method was later shown to be equivalent to the minimum residual method applied to the operator residual corresponding to a variational formulation \cite{DPG2, ChanHeuerBui-ThanhDemkowicz12}.

The derivation of the DPG method can be briefly summarized as follows: we are given a trial space $U$ and discrete trial space $U_h \subset U$, Hilbert test space $V$, and a well-posed variational problem: find $u_h\in U_h$  such that
\begin{align}
\label{variational}
&b(u_h,v) = l(v), \quad \forall v\in V, 
\end{align}
where $b(u,v)$ is a bilinear form and $l \in V'$. If we identify the variational operator $B: U\rightarrow V'$ such that $\LRa{Bu,v} \coloneqq b(u,v)$, we can write the above variational formulation as an operator equation in $V'$
\[
Bu_h = l.
\]
Seeking the minimization of the discrete residual $\frac{1}{2} \nor{Bu_h-l}^2_{V'}$, we arrive at the normal equations 
\[
\LRp{Bu_h-l,B\delta u}_{V'} = 0, \quad \forall \delta u \in U_h.  
\]
Since $V$ is Hilbert, we can introduce the Riesz map $R_V:V \rightarrow V'$, an isometric map such that $\LRa{R_V v,\cdot} = \LRp{v,\cdot}_V$. By taking the inverse of the Riesz map, we can transform the normal equations in $V'$ to a set of normal equations in $V$
\[
\LRp{R_V^{-1}\LRp{Bu_h-l},R_V^{-1}\LRp{B\delta u}}_{V} = 0, \quad \forall \delta u \in U_h.  
\]
Using the definitions of the Riesz map $R_V$ and variational operator $B$, we have that
\begin{align*}
\LRp{R_V^{-1}\LRp{Bu_h-l},R_V^{-1}\LRp{B\delta u}}_{V} = 0, \quad \forall \delta u \in U_h \\
\rightarrow \LRa{Bu_h-l,R_V^{-1}\LRp{B\delta u}}_{V} = 0, \quad \forall \delta u \in U_h \\
\rightarrow b\LRp{u_h,R_V^{-1} B\delta u} - l\LRp{R_V^{-1} B\delta u} = 0, \quad \forall \delta u \in U_h.
\end{align*}
In other words, the minimization of the operator residual associated with a given variational form corresponds directly to a Petrov-Galerkin formulation, where test functions are determined under the so-called \textit{trial-to-test} operator $R_V^{-1}B$, which, given functions in the discrete trial space $U_h$,  returns back corresponding test functions which span the discrete test space.

Solutions $u$ to the above problem enjoy the following properties: 
\begin{enumerate}
\item If we define the energy norm 
\[
\nor{u}_E = \nor{Bu}_{V'} = \sup_{v\in V\setminus\{0\}}\frac{b(u,v)}{\nor{v}_V},
\]
then DPG is optimal in $\nor{u}_E$.  This norm is induced through the energy inner product
\[
\LRp{v,w}_E = \LRp{Bv,Bw}_{V'} = \LRp{R_V^{-1}Bv, R_V^{-1}Bw}_V.
\]
\item The exact energy error $\nor{u-u_h}_E$ is computable without knowing $u$ explicitly, owing to the fact that
\[
\nor{u-u_h}_E = \nor{B(u-u_h)}_{V'} = \nor{l-Bu_h}_{V'} = \sup_{v\in V\setminus\{0\}}\frac{l(v)-b(u,v)}{\nor{v}_V},
\]
which can be computed as $\nor{e}_V$, where the \textit{error representation function} $e$ is determined as the solution to 
\[
\LRp{e,\delta v}_V = l(\delta v)-b(u,\delta v), \quad \delta v\in V.
\]
This property is simply a result of the minimum residual nature of DPG; the energy error is simply the residual, measured in the proper norm. 
\item Let $\{u_i\}_{i = 0}^{i = N}$ be discrete trial functions and $v_{u_i}$ be the corresponding optimal test functions (solutions of Problem (\ref{aux})).  The discrete system resulting from DPG is symmetric positive-definite; this can be seen from the identity
\[
b(u_i,v_{u_j}) = \LRp{v_{u_i},v_{u_j}}_V = \LRp{v_{u_j},v_{u_i}}_V = b(u_j,v_{u_i}).
\]
Positive-definiteness is again a natural consequence of the minimum-residual nature of DPG.  
\end{enumerate}

\subsection{Broken test spaces}

In practice, it can be very difficult computationally to determine a set of optimal test functions.  The inverse of the Riesz map $R_V^{-1} B\delta u$ can be computed by solving the auxiliary variational problem
\begin{equation}
\label{aux}
(v,\delta v)_V = b\LRp{\delta u,\delta v}, \quad \forall \delta v\in V.  
\end{equation}
However, for a conforming discretization, this amounts to a global problem.  For example, if $V = H^1(\Omega)$ and $\LRp{\cdot, \cdot }_V$ corresponded to the $H^1$ Sobolev inner product  $\LRp{v,\delta v}_{L^2(\Omega)} + \LRp{\grad v,\grad \delta v}_{L^2(\Omega)}$, solving for a single optimal test function would be computationally equivalent to the solution of the Poisson equation over the entire domain.  However, we can alleviate this issue by choosing for $V$ a \textit{broken} test space.

We begin by assuming we are given some triangulation $\Omega_h$ of $\Omega$ such that $\Omega = \bigoplus_{K\in \Omega_h}K$.  We can adopt for $V$ the broken test space
\[
V(\Oh) = \bigoplus_{K\in \Omega_h}V(K).
\]
If we further assume that the inner product and norm on $V$ are localizable\footnote{A localizable norm $\nor{v}_{V(\Oh)}$ can be written in the form 
$$\nor{v}_{V(\Oh)} = \sum_{K\in\Oh} \nor{v}_{V(K)},$$ where $\nor{v}_{V(K)}$ is a norm over the element $K$.}, the solution to \eqref{aux} (corresponding to the inversion of the Riesz map) can now be determined locally on each element.  In practice, these solutions are approximated by solving (\ref{aux}) over a locally conforming space $V_h$ over each element $K$, where $\dim(V_h) > \dim(U_h)$ elementwise.  Typically, if $U_h$ is of order $p$, $V_h$ is chosen to be the corresponding $H^1$ or $H({\rm div})$ conforming space of order $p + \del p$, where $\del p \geq n$, the spatial dimension \cite{practicalDPG}.  For this reason, we refer to $V_h$ as the \textit{enriched} space.  We refer to this approximated test space under $V_h$ as the approximate optimal test space $V_{{\rm opt},h}$.  
\begin{remark}
All properties of DPG carry over if $V=V_h$, the discrete inner product space, and if Problem~(\ref{aux}) is well-approximated under $V_h$, then the solution to our variational problem (\ref{variational}) over the approximate optimal test space will deliver very similar results to the solution of (\ref{variational}) over the exact optimal test space.  
\end{remark}

\subsection{The ultra-weak variational formulation}

Up to now, we've left our specific variational formulation unspecified.  For several reasons, we've chosen to use the \textit{ultra-weak formulation}, which results from writing the PDE as a first-order system, integrating all derivatives by parts, and identifying boundary traces as independent unknowns.  The formulation is given as follows: let $$\Gh = \bigcup_{K\in \Oh} \partial K$$ be the \textit{mesh skeleton}, comprising of the union of all element boundaries.  Let $Au=f$ represents the strong form of a first-order system of PDEs; the ultra-weak formulation is then
\begin{equation}
\label{variationalProblem}
b\LRp{\LRp{u,\uh},v} = \LRp{u,A_h^*v}_{\Oh} + \sum_{K \in \Oh} \LRa{\uh,v}_{\partial K} = \LRp{f,v}_{\Omega},
\end{equation}
where $u \in L^2{\Omega}$, and $\uh \in \gamma(D(A))$, the trace space of the domain of the strong operator $A$.  $A_h^*$ is the formal adjoint of $A$ acting elementwise on a broken test space, and $\LRa{\uh,v}_{\partial K}$ is the duality pairing between the trace unknown $\uh$ and the trace of the test function $v$ on an element boundary.  Under proper regularity assumptions, we have that
$$\sum_{K \in \Oh} \LRa{\uh,v}_{\partial K} = \LRa{\uh,\jump{v}}_{\Gh}.$$

Combining this variational formulation with the concept of DPG optimal test functions yields a conforming method for $u$, the field variable, and $\uh$, the trace variable. Optimal test functions are all determined locally, and optimal test functions corresponding to traces defined on the element boundary $\partial K$ have support over each element adjacent to $\partial K$.  The degrees of freedom are thus coupled together globally through the optimal test functions for trace variables.

\begin{remark}
\label{remarkDahmen}
It is possible to apply the same minimum-residual method behind DPG in an alternative fashion by casting the normal equations as a saddle-point problem, which is the approach taken by Dahmen et al.\ in \cite{DahmenVariationalStabilization}. This method requires the solution of a larger linear system than the DPG method, but can be directly applied to arbitrary variational formulations and conforming discretizations of $V$.  
\end{remark}

\subsection{The graph test norm}

Up to now, the inner product $\LRp{v,\delta v}_V$ (which induces a norm on $V$) has not been specified.  Here, we derive two important ``canonical" test norms for the ultra-weak variational formulation.  We begin first with the canonical norm in $U$. Since $\uh \in \gamma\LRp{D\LRp{A}}$, the standard norm for $\uh$ is the so-called minimum energy extension norm defined as
\begin{equation*}
\|\widehat{u}\| = \inf_{w\in D\LRp{A},
  \left.w\right|_{\Gh}=\widehat{u}} \|w\|_{D\LRp{A}}.
\end{equation*}
The canonical norm for the group variable $\LRp{u,\uh}$ is then given by
\[
\|\left(u,\widehat{u}\right)\|_U^2 = \|u\|^2_{\L} + \|\widehat{u}\|^2.
\]
It has been shown in \cite{Bui-ThanhDemkowiczGhattas11a, ChanHeuerBui-ThanhDemkowicz12} that the canonical norm $\nor{\LRp{u,\uh}}_U$ in $U$ induces (generates) the norm $\nor{v}_{V,U}$ in $V$.

The canonical norm $\nor{\LRp{u,\uh}}_U$ in $U$ provides an optimal balance between the standard norms on the field $u$ and the flux $\uh$ \cite{DPG4}. As a result, if the induced norm $\nor{v}_{V,U}$ (namely, the optimal test norm) is used to compute optimal test functions in (\ref{aux}), the finite element error in the canonical norm is the best in the norm $\nor{u}_{U}$. Unfortunately, the optimal test norm is non-localizable due to the presence of the jump term $\jump{v}$. Since the jump terms couple elements together, the evaluation of the jump terms requires contributions from all the elements in the mesh. 
Consequently, solving for an optimal test function amounts to inverting the Riesz map over the entire mesh $\Oh$, making the optimal test norm impractical.

On the other hand, since $v \in D\LRp{A^*_h}$, we can use as a test norm for $v$ the broken graph test norm: 
\[
\nor{v}_V^2 =  \|A_h^*v\|_{\L}^2 + \nor{v}_{\L}^2.
\]
This norm is a localization of $\nor{v}_{V,U}$ to allow for the solution of optimal test functions on an element-by-element basis, and can be considered to be the canonical norm on $V$.  In the DPG literature \cite{DPG4}, $\nor{v}_{V,U}$ is known as the {\em optimal test norm}, while $\nor{v}_{V}$ is known as the quasi-optimal or {\em graph} test norm.  The DPG method was first shown to be well-posed under the graph test norm for the Poisson and convection-diffusion equations in \cite{analysisDPG}.  This analysis was later extended to the large class of Friedrichs' systems in \cite{Bui-ThanhDemkowiczGhattas11b}.  %Using variants of the quasi-optimal test norm, near-optimal results have also been obtained in the context of elasticity \cite{DPGElas} and the linear Stokes equations \cite{stokesDPG}. Numerical results also suggest that the DPG method may provide a ``pollution-free" method without phase error for the Helmholtz equation \cite{DPG4}.

\section{Globally conforming optimal test functions}

Though DPG test functions are determined locally, on an element-by-element basis, it is the global properties of the test space that determine the performance of the method.  We motivate this statement first by deriving global conditions for the test space under which the ultra-weak variational formulation recovers an $L^2$ optimal solution.

@@

Under the ultra-weak formulation, all boundary conditions are treated as natural boundary conditions.  For boundary data $\uh_0$ enforced on $\Gamma_0 \subset \Gamma$, our corresponding variational formulation is 
\begin{equation*}
b\LRp{\LRp{u,\uh},v} = \LRp{u,A_h^*v}_{\Omega} + \LRa{\uh,\jump{v}}_{\Gh \setminus \Gamma_0} = \LRp{f,v}_{\Omega} - \LRa{\widehat{u}_0,v}_{\Gamma_0}.
\end{equation*}
Suppose now that our test functions come from a \textit{conforming} space - in other words, that
\[
%\LRa{\uh,\jump{v}}_{\Gh^0} = 0, \quad \forall \uh \in \gamma(D(A))
\LRa{\uh,\jump{v}}_{\Gh\setminus \Gamma_0} = 0, \quad \forall \uh \in \gamma(D(A))
\]
%is satisfied over the internal skeleton $\Gh^0 \coloneqq \Gh\setminus\Gamma$.  
is satisfied over $\Gh\setminus \Gamma_0$.  
In this case, the inter-element boundary terms vanish, $A^*_h v$ becomes $A^*v$, and the variational formulation becomes
\begin{align*}
%b\LRp{\LRp{u,\uh},v} = \LRp{u,A^*v}_{\Omega} + \LRa{\uh,\jump{v}}_{\Gamma \setminus \Gamma_0} = \LRp{f,v}_{\Omega} - \LRa{\widehat{u}_0,\jump{v}}_{\Gamma_0}.
b\LRp{\LRp{u,\uh},v} = \LRp{u,A^*v}_{\Omega} = \LRp{f,v}_{\Omega} - \LRa{\widehat{u}_0,\jump{v}}_{\Gamma_0}.
\end{align*}
Let $\bar{V} = \{\bar{v}\in V: \LRa{\uh,\jump{\bar{v}}}_{\Gh\setminus\Gamma_0} = 0, \forall \LRp{0,\uh}\in U\}$ be the conforming subspace of $V$.  If we seek $u_h \in U_h$ and if our discrete test space is chosen to be 
\[
\bar{V}_{\rm opt} = \{\bar{v}_{\delta u} \in \bar{V}: A^*\bar{v}_{\delta u} = \delta u, \forall \delta u \in U_h\},
\]
%\[
%\bar{V}_{\rm opt} = \{\bar{v}_{\delta u} \in \bar{V}: \delta u \in U_h\},
%\]
%where each test function $v_{\delta u}$ satisfies the following adjoint equation 
%\begin{align*}
%A^*v_{\delta u} &= \delta u, \\
%\left.v_{\delta u}\right |_{\Gamma\setminus\Gamma_0} &= 0,
%\end{align*}
then our variational problem reduces to
\[
\LRp{u_h,\delta u}_{\Omega} = \LRp{f,v}_{\Omega} = \LRp{u,\delta u}_{\Omega}, \quad \forall \delta u \in U_h
\]
and the solution $u_h$ is the $L^2$ projection of the exact solution $u$ onto $U_h$.

\subsection{Globally conforming DPG test spaces}

As noted above, $L^2$ optimality is achieved when the test space possesses certain \textit{global} properties.  An obvious question concerning DPG is how much non-local information can be gleaned from locally generated test spaces.\footnote{Global properties of the test norm are related to the adjoint equation, as shown in \cite{DPGrobustness, ChanHeuerBui-ThanhDemkowicz12}.  However, global properties of the test space have not been explored specifically in these prior works.}

First, we define the idea of \textit{weakly} conforming spaces.  For discrete trial variables $\LRp{u,\uh} \in U_h$, we denote a test function as weakly conforming if
\[
b\LRp{\LRp{0,\uh},v} = \LRa{\uh,\jump{v}}_{\Gh\setminus \Gamma_0} = 0, \quad \forall \LRp{0,\uh} \in U_h.
\]
In other words, a test function is weakly conforming if its inter-element jumps are orthogonal to all trace functions in the discrete space $U_h$.  We refer to the space of these test functions as $\tilde{V} = \{\tilde{v}\in V: \LRa{\uh,\jump{\tilde{v}}}_{\Gh\setminus\Gamma_0} = 0, \forall \LRp{0,\uh_h}\in U_h\}$.

Let $V_{\rm opt} = \{v_{\delta u} \in V: v_{\delta u}= R_V^{-1}B\delta u, \delta u \in U_h\}$ be the space of locally determined DPG test functions, and let 
\[
\tilde{V}_{\rm opt} = \{\tilde{v}_{\delta u} \in \tilde{V}: \tilde{v}_{\delta u} = R_{V}^{-1}B\delta u, \delta u \in U_h\}
\]
be the space of optimal test functions determined \textit{globally} over the weakly conforming space $\tilde{V}$.  In other words, the optimal test functions that span $\tilde{V}_{\rm opt}$ are the result of the inversion of the Riesz operator \textit{globally}, over the entire mesh (the idea is not new - the concept of globally optimal test functions was first introduced in \cite{DPG3pub} to prove mesh-independence).  We have the following lemma concerning the test space spanned by the locally computed optimal test functions of DPG: 
\begin{lemma}
\label{lemma1}
$\tilde{V}_{\rm opt} \subseteq V_{\rm opt}$.
\end{lemma}
\begin{proof}
We note first that, because $V_{\rm opt} \subseteq V$, we can orthogonally decompose $V = V_{\rm opt} \bigoplus V_{\rm opt}^\perp$, where, for any $v_{\rm opt}\in V_{\rm opt}$, $\LRp{v_{\rm opt},v^\perp}_V = 0$ for all $v^\perp \in V_{\rm opt}^\perp$.  Let us now choose a globally conforming test function $\tilde{v}\in \tilde{V}_{\rm opt}$.  Since $\tilde{V}_{\rm opt} \subseteq V$, we can decompose $\tilde{v} = v_{\rm opt} + v_{\rm opt}^\perp$.  Demonstrating that $v_{\rm opt}^\perp = 0$ proves the lemma.

Let $v_{\uh} \in V_{\rm opt}$ be an optimal test function corresponding to the flux variable $\uh$.  We can use the fact that $v_{\rm opt}^\perp \in V$ to substitute it into the definition of $v_{\uh}$.  By definition, 
\[
\LRp{v_{\uh},v_{\rm opt}^\perp}_{V(K)} = \LRa{\uh,v_{\rm opt}^\perp}_{\partial K} = 0,
\]
where $\LRp{\cdot,\cdot}_{V(K)}$ denotes an element-wise inner product, and $\LRa{\cdot,\cdot}_{\partial K}$ denotes the duality pairing between $\uh$ and $v$ over the boundary $\partial K$.  Summing up over all $K$, we have
\[
\LRp{v_{\uh},v_{\rm opt}^\perp}_{V(\Oh)} = \LRa{\uh,\jump{v_{\rm opt}^\perp}}_{\Gh} = 0,
\]
Thus, we can conclude that $v_{\rm opt}^\perp \in \tilde{V}$ is weakly conforming.  Then, by definition of the weakly conforming optimal test space $\tilde{V}_{\rm opt}$, for a conforming optimal test function $\tilde{v}_u$ corresponding to a field variable $u$, we have that
\[
\LRp{\tilde{v}_u,v_{\rm opt}^\perp}_{V(\Oh)} = \LRp{u,A^*_h v_{\rm opt}^\perp}_{L^2(\Omega)} = \LRp{v_u, v_{\rm opt}^\perp}_{V(K)} = \LRp{v_u, v_{\rm opt}^\perp}_{V(\Oh)} = 0,
\]
where $v_u$ is a non-conforming locally determined test function.  The above orthogonality conditions imply that, for the globally conforming test space $\tilde{V}_{\rm opt}$, 
\[
0 = \LRp{\tilde{v},v_{\rm opt}^\perp}_V = \LRp{v_{\rm opt} + v_{\rm opt}^\perp,v_{\rm opt}^\perp}_V = \LRp{v_{\rm opt}^\perp,v_{\rm opt}^\perp}_V = \nor{v_{\rm opt}^\perp}_V^2.
\]
\end{proof}

We note that Lemma~\ref{lemma1} still holds in the case where optimal test functions spanning $V_{\rm opt}$ are approximated using $V_h$, the enriched space, so long as $V_h$ is a closed subspace of $V$.  In other words, the space of globally optimal test functions approximated using a weakly conforming enriched space $\tilde{V}_h$ is contained within the approximate optimal test space $V_{{\rm opt},h}$.

Lemma~\ref{lemma1} has an immediate consequence concerning DPG solutions under weakly conforming test spaces: 
\begin{lemma}
\label{remark1}
Let $\tilde{u}$ be the field component of the DPG solution under a weakly globally conforming optimal test space, and let $u$ be the field component of the standard DPG solution.  Assuming both problems are uniquely solvable, $u = \tilde{u}$.  
\end{lemma}
\begin{proof}
This can be shown by taking the variational problem (\ref{variational}) for $u$ and $\widehat{u}$; since $\tilde{V}_{\rm opt} \subseteq V_{\rm opt}$, we can substitute in for $v$ the weakly conforming optimal test functions spanning $\tilde{V}_{\rm opt}$. Doing so reduces (\ref{variational}) to the problem for $\tilde{u}$.   
\end{proof}

\subsection{The graph test norm and globally optimal test functions}

The concept of $L^2$ optimal test functions is connected to DPG through a variant of the graph test norm; if we take for our test norm on $V$
\[
\nor{v}_V^2 = \nor{A_h^*v}_{L^2{\Omega}}^2 + \delta \nor{v}_{L^2{\Omega}}^2,
\]
and assume boundedness below of the bilinear operator $B$ (see \cite{DPG4} for details), $\nor{v}_V$ remains a norm in the limit as $\delta \rightarrow 0$.  In this limit, the auxiliary problem \eqref{aux} for optimal test functions corresponding to the field variable $u$ reduces to
\[
\LRp{A^*_h v_u, A^*_h \delta v}_{V(K)} = \LRp{u,A^*_h \delta v}_{L^2(K)}, \quad \forall \delta v \in V(K).
\]
Recalling Lemma \ref{lemma1}, this implies that weakly conforming optimal test functions satisfy
\[
\LRp{A^*_h v_u, A^*_h \delta v}_{V(\Omega)} = \LRp{u,A^*_h \delta v}_{L^2(\Omega)}, \quad \forall \delta v \in V(\Omega),
\]
which yields a weakly conforming test subspace that can be characterized as
\[
\tilde{V}_{\rm opt} = \{\tilde{v}_{\delta u} \in \tilde{V}: A^*_h \tilde{v}_{\delta u} = {\delta u}, {\delta u} \in U_h\}.
\]
In other words, the DPG test space in the limit as $\delta \rightarrow 0$ contains a weakly conforming approximation to the $L^2$-optimal test space $\bar{V}_{\rm opt}$.  The approximation is due to the fact that strong conformity is not guaranteed; let $\LRp{u,\uh}$ and $\LRp{u_h,\uh_h}$ be the exact and approximate solutions respectively to the variational problem \eqref{variationalProblem} with $V = \tilde{V}_{\rm opt}$.  We then have globally that
\[
b\LRp{\LRp{u-u_h,\uh-\uh_h},\tilde{v}} = \LRp{u-u_h,A^*_h \tilde{v}}_{\Omega} + \LRa{\uh-\uh_h,\jump{\tilde{v}}}_{\Gh} = 0, \quad \forall \tilde{v}\in \tilde{V}_{\rm opt}.
\]
If the exact flux $\uh$ is exactly representable by the approximate flux $\uh_h$, we have that 
\[
\LRp{u_h-u,\delta u}_{\Oh} = 0
\]
for all $\delta u \in U_h$, which we recognize as the condition under which the approximate solution $u_h$ is indeed the best-$L^2$ approximation of the true solution $u$ in $U_h$.

%\textcolor{red}{Flesh this out in more detail - where does the approximation of jump terms come in?  I.e. the proper norm in which to measure flux error is the minimum energy extension norm, but this introduces derivatives, which should introduce a $\epsilon^{-1}$ factor into the norm!  Why is this not observed for convection-diffusion?!?}

\section{The convection-diffusion problem}

The weakly conforming properties of DPG test spaces proven in Lemma~\ref{lemma1} motivate the question whether the global nature of DPG can be exploited to develop a more efficient method.  We discuss one particular application of this to the convection-diffusion problem for convection-dominated regimes.

The convection-diffusion problem models the concentration $u$ of some solvent in a convection field $\beta$ subject to diffusion of some magnitude $\epsilon$.  We decompose the boundary as follows:
\begin{align*}
\Gamma_{\rm in} &\coloneqq \{x\in \Gamma; \beta_n(x) < 0\}, \quad {\rm
(inflow)}\\ 
\Gamma_{\rm out} &\coloneqq \{x\in \Gamma; \beta_n(x) > 0\},
\quad {\rm (outflow)}\\
\Gamma_{0} &\coloneqq \{x\in \Gamma;\beta_n(x) = 0\},
\end{align*}
where $\beta_n \coloneqq \beta \cdot n$.

The problem is typically written in second-order form as
\begin{align*}
\div\LRp{\beta u - \epsilon \grad u} &= f\\
u &= u_{0}, \quad \text{on } \Gamma.
\end{align*}
The problem is especially of interest as a prototypical problem for boundary layer problems in computational fluid dynamics, as solutions to the convection-diffusion problem can, for certain boundary conditions, develop strong boundary layers of width $\epsilon$ at the outflow boundary $\Gamma_{\rm out}$.

\subsection{The graph test norm for the convection-diffusion problem}

We begin first with the canonical norm for DPG, the graph test norm.  Under the ultra-weak variational formulation, the graph test norm for convection-diffusion is
\[
\nor{v}_{V_{\rm graph}}^2 = \nor{\div\tau - \beta \cdot \grad v}_{L^2(\Omega)}^2 + \nor{{\epsilon}^{-1} \tau -  \grad v}_{L^2(\Omega)}^2 + \nor{v}_{L^2(\Omega)}^2
\]
where the $L^2$ norm of $\tau$ is neglected as it is not required to preserve positive-definiteness of the norm, and $\grad$ and $\div$ are understood to act elementwise.

Introduce the graph test norm here, as well as alternative test norms and problems of robustness \cite{DPGrobustness}, Denote norms as 
\begin{itemize}
\item $\nor{\cdot}_{V_{\rm graph}}$
\item $\nor{\cdot}_{V_{\rm weight}}$ ??
\end{itemize}

\subsection{Localization and boundary layers under the graph test norm}

The graph test norm typically delivers near-optimal results - for problems of wave propagation, elasticity, and Stokes flow, DPG under the graph test norm delivers solutions nearly indistinguishable from the $L^2$ projections of the exact solution \cite{DPG4, DPGElas, stokesDPG}.  However, when applied to the convection-diffusion problem, the graph test norm performed very poorly.

It was determined that the optimal test functions generated under the graph test norm developed strong boundary layers of width $\epsilon$.  Figure~\ref{fig:optTestBoundary} shows the result of 2D numerical experiments\footnote{All figures of optimal test functions are produced using the FEniCS codebase \cite{LoggMardalEtAl2012a}.} where a fine mesh was used to resolve an optimal test function resulting from (\ref{aux}), demonstrating the presence of strong boundary layers at the element inflow boundary $\partial K_{\rm in}$.

\begin{figure}[!h]
\centering
\subfigure[$v$]{\includegraphics[width=5cm]{figs/u_graph_v_K.pdf}}
\subfigure[$\tau_x$]{\includegraphics[width=5cm]{figs/u_graph_tau_x_K.pdf}}
\subfigure[$\tau_y$]{\includegraphics[width=5cm]{figs/u_graph_tau_y_K.pdf}}
\caption{$v$ and $\tau$ components of the 2D optimal test functions corresponding to the basis function $u=1$ on the reference element for $\epsilon = 0.01$. The solution has been obtained using a fine $128\times128$ mesh of triangles, with $p = 3$. }
\label{fig:optTestBoundary}
\end{figure}

Under the graph norm for convection-diffusion, Problem~(\ref{aux}) over a single element is given as
\begin{align*}
\LRp{\LRp{v,\tau},\LRp{\delta v,\delta \tau}}_{V_{\rm graph}(K)} &= \LRp{u, \Grad_h\cdot \tau - \beta\cdot \Grad_h v}_{K} + \LRp{\sigma, \frac{1}{\epsilon}\tau + \Grad_h v}_{(K)} \\
&+ \LRa{\uh,\tau_n}_{\partial K} + \LRa{\fnh,v}_{\partial K}
\end{align*}
where $\LRp{\LRp{v,\tau},\LRp{\delta v,\delta \tau}}_{V_{\rm graph}(K)}$ is 
\begin{align*}
\LRp{\LRp{v,\tau},\LRp{\delta v,\delta \tau}}_{V_{\rm graph}(K)} =& \LRp{\div\tau - \beta \cdot \grad v,\div\delta \tau - \beta \cdot \grad \delta v}_{L^2(K)} \\
&+ \LRp{{\epsilon}^{-1} \tau -  \grad v,{\epsilon}^{-1} \delta \tau -  \grad \delta v}_{L^2(K)} + \LRp{v,\delta v}_{L^2(K)} 
\end{align*}

We can transform the above problem to the reference element.  Applying this simple scaling argument shows that, for elements of size $h$, we can expect a boundary layer of width $h/\epsilon$ relative to a unit domain.\footnote{This is assuming that the parameter $\epsilon$ dictates the width of expected boundary layers on a unit domain. The strong form of the above variational problem corresponds to a reaction-diffusion system, for which we expect this assumption to hold.  Numerical experiments also appear to confirm that the boundary layer is of width $O(\epsilon)$.}  In other words, the strength of the boundary layer is proportional to the element Peclet number ${\rm Pe} = h/\epsilon$.  For severely underresolved meshes where $h \gg \epsilon$, this makes the approximation of optimal test functions using a simple $p$-enriched space very difficult, though subgrid meshes have been used to resolve such test functions with some success in \cite{DBLP:journals/procedia/NiemiCC11}.  The construction of the alternative test norm $\nor{\cdot}_{V_{\rm weight}}$ was motivated by the computational difficulty in applying the graph test norm to heavily convection-dominated regimes where $\epsilon \ll 1$.
